---
title: "The Ridge Model"
author: "Yujun Pan"
output:
  html_document:
    self_contained: yes
    toc: true
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The Ridge Regression vignette}
-->

# Ridge Regression

Ridge regression is a shrinkage method that imposes a penalty on the magnitude of the coefficients to regularize the model. In ridge regression, coefficients get shrunken towards 0 and standard errors of coefficients are much lower. This assignment asks us to build our own ridge regression and choose the optimal hyperparameter, lambda. 

### load datasets

```{r}
library(bis557)
data(ridge_test)
data(ridge_train)
```

### ridge class
```{r}
predict.ridge_reg = function(object,...) {
  newdata = list(...)[[1]]
  m = model.matrix(object$form, newdata)
  m %*% object$coefficients
}

```

### train data and calculate out of sample MSE
```{r}
lambda_seq = seq(0.01, 1000, by=0.1)
MSEs = rep(NA, length(lambda_seq))
for (i in 1:length(lambda_seq)){
  fit = ridge_reg(y ~., lambda_seq[i], ridge_train)
  error = ridge_test$y - predict(fit, ridge_test)
  MSEs[i] = mean(error^2)
}
```

### plot
```{r}
plot(log(lambda_seq), MSEs, xlab = "log(lambda)", ylab = "Out-of-Sample MSE")
```

### show lambda with smallest MSE
```{r}
lambda_seq[which.min(MSEs)]
```